---
title: "Quantium Project"
author: "IghodaloStephanie"
output: word_document
date: "2026-05-28"
---
My Key Findings:

Young Singles/Couples generated the highest chip sales.
Mainstream customers purchased the largest quantity.
175g pack sizes were most popular.
Brand X dominated premium segments.

My Recommendation:
The company should target Mainstream Young Singles/Couples.
Increase promotions on popular pack sizes.
Place premium brands near high-traffic areas.


```{r}
Behaviour_data <- read.csv("QVI_purchase_behaviour.csv")

head(Behaviour_data)
```

package installation


install.packages("tidyverse", repos = "https://r-project.org")
install.packages("janitor")
install.packages("lubridate")
install.packages("stringr")

Import excel file(install package first)


install.packages("readxl")



```{r}
library(readxl)
Transaction_data <- read_excel("QVI_transaction_data.xlsx")
 
head(Transaction_data)
```
```{r}
str(Behaviour_data)
str(Transaction_data)
```

Summary statistics
```{r}
summary(Behaviour_data)
summary(Transaction_data)
```
Check for missing values
```{r}
colSums(is.na(Transaction_data))

colSums(is.na(Behaviour_data))
```
No missing values in both data

Checking for duplicates
```{r}
sum(duplicated(Transaction_data))
sum(duplicated(Behaviour_data))
```
Removing duplicates
```{r}
library(tidyverse)
Transaction_data <- distinct(Transaction_data)
sum(duplicated(Transaction_data))
```
Converting Excel numeric date to actual date
```{r}
Transaction_data$DATE <- as.Date(Transaction_data$DATE, origin = "1899-12-30")
head(Transaction_data$DATE)
```

```{r}
unique(Transaction_data$PROD_NAME)
```
Count products containing salsa and Filter out salsa products
```{r}
Transaction_data %>%
  filter(str_detect(PROD_NAME, "salsa|Salsa"))

Transaction_data <- Transaction_data %>%
  filter(!str_detect(PROD_NAME, "salsa|Salsa"))
```

```{r}
Transaction_data <- Transaction_data %>%
  mutate(PACK_SIZE = str_extract(PROD_NAME, "\\d+")) %>%
  mutate(PACK_SIZE = as.numeric(PACK_SIZE))


head(Transaction_data$PACK_SIZE)
```
```{r}
Transaction_data <- Transaction_data %>%
  mutate(BRAND = word(PROD_NAME, 1))

unique(Transaction_data$BRAND)
```
Fixing inconsistent spellings
```{r}
Transaction_data$BRAND <- recode(Transaction_data$BRAND,
                             "RED" = "RRD")

Transaction_data %>%
  ggplot(aes(x = PROD_QTY)) +
  geom_boxplot()
head(Transaction_data)
```

Customers purchasing unusually high quantities and remove outlier Customers
```{r}
Transaction_data %>%
  group_by(LYLTY_CARD_NBR) %>%
  summarise(total_qty = sum(PROD_QTY)) %>%
  arrange(desc(total_qty))

Transaction_data <- Transaction_data %>%
  filter(LYLTY_CARD_NBR != 226000)
```

Merge data

```{r}
library("tidyverse")
merged_data <- Transaction_data %>%
  left_join(Behaviour_data, by = "LYLTY_CARD_NBR")

head(merged_data)
```

merged_data <- merged_data %>%
  mutate(TOT_SALES = PROD_QTY * TOT_SALES)
  


Analyzing Sales by Customer Segment

```{r}
segment_sales <- merged_data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(
    total_sales = sum(TOT_SALES),
    total_customers = n_distinct(LYLTY_CARD_NBR),
    avg_sales = mean(TOT_SALES)
  )

segment_sales

```

Top-Selling Brands
```{r}
brand_sales <- merged_data %>%
  group_by(BRAND) %>%
  summarise(total_sales = sum(TOT_SALES)) %>%
  arrange(desc(total_sales))

brand_sales
```

Analyzing Pack Size Preferences

```{r}
pack_sales <- merged_data %>%
  group_by(PACK_SIZE) %>%
  summarise(total_sales = sum(TOT_SALES)) %>%
  arrange(desc(total_sales))

pack_sales
```

Visuals

Sales by Life Stage

```{r}
merged_data %>%
  group_by(LIFESTAGE) %>%
  summarise(total_sales = sum(TOT_SALES)) %>%
  ggplot(aes(x = reorder(LIFESTAGE, total_sales),
             y = total_sales)) +
  geom_col() +
  coord_flip()
```

Sales by Premium Segment

```{r}
merged_data %>%
  group_by(PREMIUM_CUSTOMER) %>%
  summarise(total_sales = sum(TOT_SALES)) %>%
  ggplot(aes(x = PREMIUM_CUSTOMER,
             y = total_sales)) +
  geom_col()

theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

Plot transactions over time

```{r}

transactions_over_time <- Transaction_data %>%
  group_by(DATE) %>%
  summarise(num_transactions = n())

transactions_over_time %>%
  ggplot(aes(x = DATE, y = num_transactions)) +
  geom_line() +
  labs(
    title = "Transactions Over Time",
    x = "Date",
    y = "Number of Transactions"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

Top Brands
```{r}
brand_sales %>%
  top_n(10, total_sales) %>%
  ggplot(aes(x = reorder(BRAND, total_sales),
             y = total_sales)) +
  geom_col() +
  coord_flip()
```






